Experience Report: Haskell in Computational Biology

### 2012-06-28 mrfy_experience_report.tex

\category{J.3}{Biology and genetics}{}
\category{D.3.2}{Language Classifications}{Applicative (functional) languages}
% \category{CR-number}{subcategory}{third-level}
% 
% \terms
% term1, term2
% 
% \keywords
% keyword1, keyword2

@@

The \texttt{Scored} abstraction makes it easy to implement
Viterbi's equations.
For example, to compute $V_j^{\prime M}(i)$ using the equation at the top
right of \figref{viterbi}, we call 
\mbox{\texttt{vee\char`\'} \texttt{Mat} $j$ $i$}.
The equation adds~$e^{\prime}_{M_{j}}(x_{i})$, computed
with \texttt{eScore}, to a minimum.
The sum of an~$a'$~term and a $V'_{j-1}(i-1)$ term 
is computed by function \texttt{avSum},
the terms are computed by \texttt{aScore} and \texttt{vee''}, respectively.
\smallverbatiminput{vfix}
The entire right-hand side is wrapped in \texttt{extend~Mat}.
This call performs a computation \emph{not} shown in \figref{viterbi}:
in~addition to computing the most likely BUZZWORD,
\mrfy\ also computes most likely path (sequence of states).
Function \texttt{extend} adds~$M$ to the head of that path.

>>>>

\category{J.3}{Biology and genetics}{}
\category{D.1.1}{Applicative (Functional) Programming}{}
% \category{CR-number}{subcategory}{third-level}
% 
% \terms
% term1, term2
% 
% \keywords
% keyword1, keyword2

@@

The \texttt{Scored} abstraction makes it easy to implement
Viterbi's equations.
For example, to compute $V_j^{\prime M}(i)$ using the equation at the top
right of \figref{viterbi}, we call 
\mbox{\texttt{vee\char`\'} \texttt{Mat} $j$ $i$}.
The equation adds~$e^{\prime}_{M_{j}}(x_{i})$, computed
with \texttt{eScore}, to a minimum.
The sum of an~$a'$~term and a $V'_{j-1}(i-1)$ term 
is computed by function \texttt{avSum},
the terms are computed by \texttt{aScore} and \texttt{vee''}, respectively.
\smallverbatiminput{vfix}
The entire right-hand side is wrapped in \texttt{extend~Mat}.
This call performs a computation \emph{not} shown in \figref{viterbi}:
in~addition to computing the probability of the most likely path,
\mrfy\ also computes the most likely path (sequence of states).
Function \texttt{extend} adds~$M$ to the head of that path.
